\texttt{ensmallen}: a flexible C++ library for efficient function optimization

### 2018-10-17 doc/mlsys_paper/paper.tex
combine into one figure the '6 optimizers' code snippet with the corresponding learning curves plot

The results indicate that {\tt ensmallen} with \texttt{\small EvaluateWithGradient()}
is the fastest approach.
Furthremore, the use of \texttt{\small EvaluateWithGradient()} yields considerable
speedup of almost 1.5x over the {\tt ensmallen} implementation with both the
objective and gradient implemented separately.  In addition, although the
automatic differentiation support makes it easier for the user to write their
code, it is clear that the code generated by the automatic differentiators
({\tt Calculus.jl} and {\tt ForwardDiff.jl})
is far from optimal, with runs that are slower by up to three orders of magnitude.

@@

Lastly, we show the simplicity of using different optimizers for the same task.
Using a version of {\tt LinearRegressionFunction} adapted for separable
differentiable optimizers, we run six optimizers with default parameters in
eight lines of code:

% RC: I know this is inline code... but does it work here?  If not we can put it
% in a figure.
\begin{minted}[fontsize=\footnotesize]{c++}
  LinearRegressionFunction lrf(X, y); // X and y are data

  ens::StandardSGD<>().Optimize(lrf, sgdModel);
  ens::Adam().Optimize(lrf, adamModel);
  ens::AdaGrad().Optimize(lrf, adagradModel);
  ens::SMORMS3().Optimize(lrf, smorms3Model);
  ens::SPALeRASGD().Optimize(lrf, spaleraModel);
  ens::RMSProp().Optimize(lrf, rmspropModel);
\end{minted}

This pluggability makes it easy to use {\tt ensmallen} to generate comparisons
between optimizers. In Fig.~\ref{fig:learning_curve} we use these six
lines of code to generate learning curves for each of these six optimizers on
the {\tt YearPredictionMSD} dataset from the UCI dataset
repository~\cite{ucimlrepository}. % TODO: add to bibtex
In addition, any other optimizers for separable differentiable objective
functions can be dropped into place.

\begin{figure}[h!]
\begin{minipage}[t]{1\textwidth}
\begin{minipage}{0.60\textwidth}
  \centering
  \includegraphics[width=\textwidth,height=0.45\textwidth]{experiments/learning_curves_crop.png}
\end{minipage}
\hfill
\begin{minipage}{0.38\textwidth}
  \centering
  \caption
    {
    Learning curves for six {\tt ensmallen} optimizers 
    for optimizing a linear regression function 
    on the {\tt YearPredictionMSD} dataset~\cite{ucimlrepository}
    over 5 epochs of training.
    Optimizers can be tuned for better performance.
    }
  \label{fig:learning_curve}
\end{minipage}
\end{minipage}
\end{figure}

>>>>

The results indicate that {\tt ensmallen} with \texttt{\small EvaluateWithGradient()}
is the fastest approach.
Furthremore, the use of \texttt{\small EvaluateWithGradient()} yields considerable
speedup of almost 1.5x over the {\tt ensmallen} implementation with both the
objective and gradient implemented separately.  In addition, although the
automatic differentiation support makes it easier for users to write their
code, it is clear that the code generated by the automatic differentiators
({\tt Calculus.jl} and {\tt ForwardDiff.jl})
is far from optimal, with runs that are slower by up to three orders of magnitude.

@@

Lastly, we demonstrate the easy pluggability in {\tt ensmallen}
for using various optimizers on the same task.
Using a version of \texttt{\small LinearRegressionFunction} from Sec.~\ref{sec:linreg_example}
adapted for separable differentiable optimizers,
we run six optimizers with default parameters in just 7 lines of code,
as shown in Fig.~\ref{fig:learning_curve}(a).
Applying these optimizers to the {\tt YearPredictionMSD}
dataset from the UCI repository~\cite{ucimlrepository} % TODO: add to bibtex
yields the learning curves shown in Fig.~\ref{fig:learning_curve}(b).
Any other optimizers for separable differentiable objective
functions can be dropped into place in the same manner.

% This facilitates the evaluation of various optimizers
% for user-defined objective functions.

\begin{figure}[h!]
\begin{minipage}[t]{1\textwidth}
\begin{adjustbox}{scale={0.90}{0.90}}
\begin{minipage}{0.50\textwidth}
\centering
\begin{minted}[fontsize=\footnotesize]{c++}
// X and y are data
LinearRegressionFunction lrf(X, y);

ens::StandardSGD<>().Optimize(lrf, sgdModel);
ens::Adam().Optimize(lrf, adamModel);
ens::AdaGrad().Optimize(lrf, adagradModel);
ens::SMORMS3().Optimize(lrf, smorms3Model);
ens::SPALeRASGD().Optimize(lrf, spaleraModel);
ens::RMSProp().Optimize(lrf, rmspropModel);
\end{minted}
\end{minipage}
\end{adjustbox}
\hfill
\begin{minipage}{0.49\textwidth}
  \centering
  \includegraphics[width=\textwidth,height=0.45\textwidth]{experiments/learning_curves_crop.png}
\end{minipage}
\end{minipage}
\begin{minipage}[t]{1\textwidth}
\begin{minipage}{0.50\textwidth}
\centering
(a)
\end{minipage}
\hfill
\begin{minipage}{0.49\textwidth}
\centering
(b)
\end{minipage}
\end{minipage}
\caption
  {
  (a): example code for using 
  six {\tt ensmallen} optimizers with default parameters
  to optimize a linear regression function;
  (b): corresponding learning curves
  on the {\tt YearPredictionMSD} dataset~\cite{ucimlrepository}
  over 5 epochs of training.
  The optimizers can be tuned for better performance.
  }
\label{fig:learning_curve}
\end{figure}
